---
title: "Quick start guide to TheiaR package"
author: "Xavier Laviron"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{"Quick start guide to TheiaR package"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## TheiaR package

The TheiaR package intends to provide a clean interface to download and manage
data issued by [Theia](https://theia.cnes.fr/atdistrib/rocket/#/home).


## Create a TheiaAuth object

To search and download data fro Theia, you need to register to their
[website](https://sso.theia-land.fr/theia/register/register.xhtml).
Once you have credentials, store them in simple text file, let's say `auth.txt`,
with only two lines: your login and your password.

Then create a `TheiaAuth` object:

```{r, eval=F}
TheiaAuth$new(auth.file = "auth.txt")
```

You will use this object as identification for all queries needing
identification (downloads). It will generate a token during its creation, and
automatically generate a new token when needed (tokens are only valid for 2
hours).


## Searching Theia database

You can use the `TheiaQuery` framework to manage searches.
